---
title: "Linear Models"
author: "Billy Lozowski"
date: '`r Sys.Date()`'
output: 
  pdf_document:
    toc: true
  html_document:
    toc: true
  md_document:
      variant: gfm
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Read in the data called “PlantEmergence.csv” using a relative file path and load the
following libraries. tidyverse, lme4, emmeans, multcomp, and multcompView. Turn the
Treatment, DaysAfterPlanting and Rep into factors using the function as.factor

```{r packages, warning=FALSE, message=FALSE}
library(tidyverse)
library(lme4)
library(emmeans)
library(multcomp)
library(multcompView)
```

```{r data}
plant.emergence <- read.csv("data/PlantEmergence.csv")

plant.emergence$Treatment = as.factor(plant.emergence$Treatment)
plant.emergence$DaysAfterPlanting = as.factor(plant.emergence$DaysAfterPlanting)
plant.emergence$Rep = as.factor(plant.emergence$Rep)

```

2. Fit a linear model to predict Emergence using Treatment and DaysAfterPlanting along with the interaction. Provide the summary of the linear model and ANOVA results.

Treatment had a strong effect on Emergence (F = 307.95, p < .001). On average, emergence occurred after 182.3 days (I'm unsure if these are the appropriate units?!), however, after Treatment2, this was 136.5 days earlier.

```{r interaction model}
lm_model <- lm(Emergence ~ Treatment + DaysAfterPlanting + Treatment:DaysAfterPlanting,
               data = plant.emergence)
summary(lm_model)
anova(lm_model)
```

3. Based on the results of the linear model in question 2, do you need to fit the
interaction term? Provide a simplified linear model without the interaction term but still
testing both main effects. Provide the summary and ANOVA results. Then, interpret the
intercept and the coefficient for Treatment 2.

An interaction term was not necessary, since none exhibited statistical or practical significance within the linear model.

According to the simplified linear model, Treatment still has a strong effect on Emergence (all but Treatments 4, 7, and 9 were p < .05). However, this model also revealed that DaysAfterPlanting also had a strong effect (all ps < .001).

```{r sans interaction model}
lm_model1 <- lm(Emergence ~ Treatment + DaysAfterPlanting, 
                data = plant.emergence)
summary(lm_model1)
anova(lm_model1)
```
4. Calculate the least square means for Treatment using the emmeans package and
perform a Tukey separation with the compact letter display using the cld function.
Interpret the results

According to visualisations and pairwise comparisons, the effect for Treatment2 was statistically different from all other treatment types. With a mean Emergence of 55.6 (units), Treatment2 was roughly 135 units sooner than all others (all ps < .001).


```{r least squares}
lsmeans <- emmeans(lm_model1, ~Treatment, data = plant.emergence)
ggplot(data = plant.emergence, aes(x = Treatment, y = Emergence, fill = Treatment)) +
  geom_boxplot()
summary(lsmeans)

results_lsmeans <- cld(lsmeans, alpha = 0.05, details = TRUE)
summary(results_lsmeans)

```

5. The provided function lets you dynamically add a linear model plus one factor from
that model and plots a bar chart with letters denoting treatment differences. Use this
model to generate the plot shown below. Explain the significance of the letters.

```{r lm function}
plot_cldbars_onefactor <- function(lm_model, factor) {
  data <- lm_model$model
  variables <- colnames(lm_model1$model)
  dependent_var <- variables[1]
  independent_var <- variables[2:length(variables)]
  
  lsmeans <- emmeans(lm_model, as.formula(paste("~", factor))) # estimate
  lsmeans
  
  results_lsmeans <- cld(lsmeans, alpha = 0.05, 
                         reversed = TRUE, 
                         details = TRUE, 
                         Letters = letters) # contrast with Tukey adjustment by default.

# Extracting the letters for the bars
  sig.diff.letters <- data.frame(results_lsmeans$emmeans[,1],
                                 str_trim(results_lsmeans$emmeans[,7]))
  colnames(sig.diff.letters) <- c(factor, "Letters")

# for plotting with letters from significance test
  ave_stand2 <- lm_model$model %>%
    group_by(!!sym(factor)) %>%
    dplyr::summarize(
      ave.emerge = mean(.data[[dependent_var]], na.rm = TRUE),
      se = sd(.data[[dependent_var]]) / sqrt(n())
      ) %>%
    left_join(sig.diff.letters, by = factor) %>%
    mutate(letter_position = ave.emerge + 10 * se)
  
  plot <- ggplot(data, aes(x = !! sym(factor), y = !! sym(dependent_var))) +
    stat_summary(fun = mean, geom = "bar") +
    stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.5) +
    ylab("Number of emerged plants") +
    geom_jitter(width = 0.02, alpha = 0.5) +
    geom_text(data = ave_stand2, aes(label = Letters, y = letter_position),
              size = 5) +
    xlab(as.character(factor)) +
    theme_classic()
  return(plot)
}

plot_cldbars_onefactor(lm_model1, "Treatment")
```

6. Generate the gfm .md file along with a .html, .docx, or .pdf. Commit, and push the
.md file to github and turn in the .html, .docx, or .pdf to Canvas. Provide me a link here to
your github.

[Coding Challenge 7](Coding Challenges/Challenge7/Coding-Challenge-7.md)
