---
title: "R Markdown in-class"
author: "Billy Lozowski"
date: "`r Sys.Date()`"
output:
  word_document:
    toc: true
  html_document:
    toc: true
    word_document: null
    pdf_document: null
  pdf_document:
    toc: true
  md_document:
      variant: gfm
---


### Question 1
a.	A YAML header specifies certain information about the markdown file like the title, output type (html/pdf etc.), and any other knitted formatting arguments.

b.	Literate programming refers to explaining what code does literally (say annotations). I assume this is so someone can follow exactly what you're trying to achieve within your code.

--------------------------------------------------------------------------------

[&rarr; Noel et al. (2022) article &larr;](https://doi.org/10.1094/PDIS-06-21-1253-RE)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo = FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(ggpubr)
library(ggrepel)

# load the data
mycotoxin <- read.csv("Data/MycotoxinData.csv", na.strings = "na")

# set colour palette for figures
cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

```

### DON (ppm) x Treatment

```{r warning=FALSE, message=FALSE}
p1 <- ggplot(mycotoxin, aes(Treatment, DON, fill = Cultivar)) +       # set plot mapping
  geom_boxplot(position = position_dodge(0.85), outlier.shape = NA) + # dodge the boxplots + hide outliers
  geom_jitter(shape = 21, width = 0.3, alpha = 0.6) +                 # set point shape, dodge width, and transparency
  scale_fill_manual(values = c(cbbPalette[(3)], cbbPalette[(4)])) +   # specify fill colours
  labs(x = "",                                                        # define labels
       y = "DON (ppm)") +
  theme_classic() +
  facet_wrap(~Cultivar)                                               # separate plots by Cultivar

P1 <- p1 + 
  stat_compare_means(method = "anova") +  
  geom_pwc(aes(group = Treatment),  # specify which means you want to compare
               method = "t.test",
               label = "{p.adj.format}{p.adj.signif}")
P1

```

### X15ADON x Treatment

```{r warning=FALSE, message=FALSE}
p3 <- ggplot(mycotoxin, aes(Treatment, X15ADON, fill = Cultivar)) +          # mapping
  geom_boxplot(position = position_dodge(0.85), outlier.shape = NA) +        # boxplot mapping
  geom_jitter(shape = 21, width = 0.3, alpha = 0.6) +                        # jitter mapping
  scale_fill_manual(values = c(cbbPalette[(3)], cbbPalette[(4)])) +          # fill colours
  labs(x = "",                                                               # labels
       y = "15ADON") +
  theme_classic() +
  facet_wrap(~Cultivar)                                                      # separate plots

P3 <- p3 +
  stat_compare_means(method = "anova") +  
  geom_pwc(aes(group = Treatment),  # specify which means you want to compare
               method = "t.test",
               label = "{p.adj.format}{p.adj.signif}")
P3

```

### Seed Mass x Treatment

```{r warning=FALSE, message=FALSE}
p4 <- ggplot(mycotoxin, aes(Treatment, MassperSeed_mg, fill = Cultivar)) +   # mapping
  geom_boxplot(position = position_dodge(0.85), outlier.shape = NA) +        # boxplot mapping
  geom_jitter(shape = 21, width = 0.3, alpha = 0.6) +                        # jitter mapping
  scale_fill_manual(values = c(cbbPalette[(3)], cbbPalette[(4)])) +          # fill colours
  labs(x = "",                                                               # labels
       y = "Seed Mass (mg)") +
  theme_classic() +
  facet_wrap(~Cultivar)                                                      # separate plots

P4 <- p4 +
  stat_compare_means(method = "anova") +  
  geom_pwc(aes(group = Treatment),  # specify which means you want to compare
               method = "t.test",
               label = "{p.adj.format}{p.adj.signif}")
P4
```

### Combined plot with 'ggarrange'

```{r warning=FALSE, message=FALSE}
# arrange plots + stats in combined figure
fig2 <- ggarrange(
  P1, P3, P4,        # select plot objects
  labels = "auto",
  nrow = 1,
  ncol = 3,
  legend = FALSE     # remove legend
)
fig2
```

[GitHub Link](https://github.com/billylozowski/PLPA_5820.git)

